The human in the loop: volunteer computing as a socio-technical system

### 2016-02-29 Alife16/paper.tex
Minor corrections and some questions for @JJ in SOA section for issue #101 (although it was closed)

Volunteer computing involves an user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \citep{david-seti:home}. However
the dual introduction of JavaScript as an universal language for the
browser and the browser as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here and whose first version
was described in \citep{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used for either unwitting
\citep{unwitting-ec,boldrin2007distributed} or volunteer 
\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
used Java \citep{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms %volunteer systems?
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance.

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer. Apolónia et al. \citep{apolonia2012enhancing} used the
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. 
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

>>>>

Volunteer computing involves a user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \citep{david-seti:home}. However
the dual introduction of JavaScript as a universal language for the
browser and the browser as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here, and whose first version
was described in \citep{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used for either unwitting
\citep{unwitting-ec,boldrin2007distributed} or volunteer 
\citep{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\citep{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
%Pablo: Desell in 2008 is not so recent
used Java \citep{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\citep{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower. %Pablo: this is useful for us because...

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms %volunteer systems?
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. %Pablo: On the contrary, in this paper, we...

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
%Pablo: I miss some definition for the concept "sociotechnical metacomputer" in the introduction
metacomputer. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. %Pablo: Add a phrase here to better link with the previous description of the work of Apolonia (it is a sudden huge jump from one topic to another)
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.
